Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-27 manuscript/ERL_global_C_SI.Rmd
add/ cite Table S1, misc cleanup

## Table S#(variable mapping))

[this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))

>>>>

## Table S1. Numbers of records by biome and age class
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
table <- read.csv("tables_figures/Biomes_n_records_variables_of_interest.csv",stringsAsFactors = FALSE, check.names = FALSE)
table_noAgeNA <- subset(table, Age != 'NA') # remove records with age unknown
table_wide <- reshape(table_noAgeNA, idvar = "Biome", timevar = "Age", direction = "wide")
colnames(table_wide)[2] <- "Mature"
colnames(table_wide)[3] <- "Young"
kable(table_wide, format = "latex", booktabs = TRUE, escape = F)%>%
  add_header_above(c(" ", "n records" = 2)) %>%
  pack_rows("Boreal climate zones", 1, 3, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
  pack_rows("Excluded climate zones", 4, 6, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
  pack_rows("Temperate climate zones", 7, 9, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
  pack_rows("Tropical climate zones", 10, 12, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
  row_spec(2, bold = T) %>%
  row_spec(7:8, bold = T) %>%
  row_spec(10, bold = T) %>%
  kable_styling( protect_latex = T, font_size = 10)
```
Sample sizes refer to data set after merging of duplicates, removal of stands with no age or history data, and removal of managed and disturbed stands. For vegetation type, "Other" refers to stands that are mixed broadleaf/ conifer or that have not been classified. Focal biomes are indicated in bold.

### 2020-09-27 manuscript/ERL_global_C_review.Rmd
add/ cite Table S1, misc cleanup

# limiting to variables and biomes of interest:
 Biomes_n_records <- read.csv("tables_figures/Biomes_n_records_variables_of_interest.csv", check.names = F)
n_focal_variable_records_eligible_for_inclusion <- sum(Biomes_n_records$n_records) # n records_eligible_for_inclusion, limited to focal variables
# # records for excluded biome groupings ? 
# n_records_BoB <- NA #Biomes_n_records$n_records[Biomes_n_records$Biome %in% "Boreal broadleaf"] # for all n, subset by Age if needed
# n_records_Bomixed <- NA
# n_records_Te_mixed <- NA

@@

n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes])
n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks])

```

@@

*Review Results/ Synthesis.*  ForC v.**XX** yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets. The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation. The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.

*Discussion.*  *NEED TO WRITE THIS!!!*

@@

We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC (Table 1).
Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)).
Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum threshold for tree census was $\le$ 10cm. 
All records were measured directly or derived from field measurements (as opposed to modeled).

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database). We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.

@@

To test for differences across mature forest biomes, the original values were analyzed via a linear mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; **REF**] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
There were enough data to run this analysis for all focal variables but **$DW_{down}$**.

@@

Of the `r nrow(ForC_measurements)` records in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1).
These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.

@@

Average C cycles for forests <100 years old are presented in Figures 8-11.  
Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs. 6-11, S1-S26; detailed below).

@@

There were only *##* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$. **(more could be added if age trends become significant after outliers are resolved)**
For three of these ($ANPP$, $ANPP_{litterfall}$, $BNPP$), the increase in C flux with age was steepest increase in tropical forests, followed by temperate and then boreal forests (Figs S#).
Similarly, $ANPP_{woody}$ displayed a steeper increase with age in temperate than boreal boreal forests (no tropical data for this variable).
In contrast, for $ANPP_{stem}$,  tropical and temperate broadleaf forests had the highest flux rates at young ages, but were surpassed by temperate conifer forests between ages 20 and 50 (Fig. S6).

@@

ForC v.XX brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs. 6-7).
In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, 7).
The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-11).
Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.

@@

## ORCID iD

>>>>

# limiting to variables and biomes of interest:
 Biomes_n_records <- read.csv("tables_figures/Biomes_n_records_variables_of_interest.csv", check.names = F)
n_focal_variable_records_eligible_for_inclusion <- sum(Biomes_n_records$n_records) # n records_eligible_for_inclusion, limited to focal variables
# # records for excluded biome groupings ? 
# n_records_BoB <- NA #Biomes_n_records$n_records[Biomes_n_records$Biome %in% "Boreal broadleaf"] # for all n, subset by Age if needed

@@

n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes])
n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks])
```

@@

*Review Results/ Synthesis.*  ForC v3.0 yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets. The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation. The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.

*Discussion.*  **NEED TO WRITE THIS!!!**

@@

We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC (Table 1).
Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables. Specifically, net ecosystem exchange (measured by eddy-covariance; **FLUXNET REF**) and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
Furthermore, for $NPP$, $ANPP$, and $ANPP_{litterfall}$, ForC variables specifying inclusion of different components were combined.
Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum threshold for tree census was $\le$ 10cm. 
All records were measured directly or derived from field measurements (as opposed to modeled).

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database). We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.
Numbers of records by biome and age class are given in Table S1.

@@

To test for differences across mature forest biomes, the original values were analyzed via a linear mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; **REF**] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
There were enough data to run this analysis for all focal variables but **$R_{auto-ag}$ and $DW_{down}$**.

@@

Of the `r nrow(ForC_measurements)` records in ForC v3.0, `r n_records` met the criteria for inclusion in this study (Fig. 1).
These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.

@@

Average C cycles for forests <100 years old are presented in Figures 8-11.  
Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).

@@

There were only *##* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$ (Table S2). **(more could be added if age trends become significant after outliers are resolved)**
For three of these ($ANPP$, $ANPP_{litterfall}$, $BNPP$), the increase in C flux with age was steepest increase in tropical forests, followed by temperate and then boreal forests (Figs S#).
Similarly, $ANPP_{woody}$ displayed a steeper increase with age in temperate than boreal boreal forests (no tropical data for this variable).
In contrast, for $ANPP_{stem}$,  tropical and temperate broadleaf forests had the highest flux rates at young ages, but were surpassed by temperate conifer forests between ages 20 and 50 (Fig. S6).

@@

ForC v3.0 brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs. 6-7).
In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, 7).
The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-11).
Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.

@@

## ORCID iD
Kristina J. Anderson-Teixeira: https://orcid.org/0000-0001-8461-9713
